---
title: "Cup Assay Updated"
author: "Devon"
date: "May 9, 2016"
output: html_document
---

Ordered Emergence Bar Graph Relative to Control[trial]

```{r}
library(scales)
library(ggplot2)
AllCup <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
DW2<-aggregate(AllCup$emergetrial, list(AllCup$Title,AllCup$Var), na.rm=TRUE, mean)
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))
serror <- (aggregate(AllCup$emergetrial, list(AllCup$Title,AllCup$Var), se))
DWA <- cbind(DW2, serror$x)
#DW <- DW[-c(1),]
colnames(DWA) <- c("trt", "var", "eme", "SE")
limits<-aes(ymax=eme+SE, ymin=eme-SE)
#GH$path<-factor(GH$path, labels=c("Control", "P soj", "Py syl", "F oxy", "F sol", "AG 2-II", "AG 4"))
#sigletters <- data.frame("a","b","c","d")yield$var<-factor(yield$var, labels=c("AG1", "AG2", "P1", "P2"))
#DW <- transform(DW, trt = reorder (trt, rdw)
y <- ggplot(DWA, aes(y=eme, 
                    #fill=sdtrt, 
                    x=factor(DWA$trt)))
y+geom_bar(position="dodge", aes(x=reorder(trt, -eme)),  stat="identity")+
facet_grid(var~.)+
geom_errorbar(limits, position = "dodge")+
labs(title="Emergence Effect of Pythium spp. Inoculation", x="Pythium spp.", y="Emergence \n (% of the NTC, Rice)")+
 theme(plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=1), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 10),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(0,120), oob=rescale_none)
```



Ordered Root Area Bar Graph Relative to Control[trial]

```{r} 
library(scales)
library(ggplot2)
AllCup <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
DW2<-aggregate(AllCup$rtareatrial, list(AllCup$Title,AllCup$Var), na.rm=TRUE, mean)
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))
serror <- (aggregate(AllCup$rtareatrial, list(AllCup$Title,AllCup$Var), se))
DWA <- cbind(DW2, serror$x)
#DW <- DW[-c(1),]
colnames(DWA) <- c("trt", "var", "rat", "SE")
limits<-aes(ymax=rat+SE, ymin=rat-SE)
#GH$path<-factor(GH$path, labels=c("Control", "P soj", "Py syl", "F oxy", "F sol", "AG 2-II", "AG 4"))
#sigletters <- data.frame("a","b","c","d")yield$var<-factor(yield$var, labels=c("AG1", "AG2", "P1", "P2"))
#DW <- transform(DW, trt = reorder (trt, rdw)
y <- ggplot(DWA, aes(y=rat, 
                    #fill=sdtrt, 
                    x=factor(DWA$trt)))
y+geom_bar(position="dodge", aes(x=reorder(trt, -rat)),  stat="identity")+
facet_grid(var~.)+
geom_errorbar(limits, position = "dodge")+
labs(title="Root Area Effect of Pythium spp. Inoculation", x="Pythium spp.", y="Root Area \n (% of the NTC, Rice)")+
  theme(plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=1), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 10),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(0,120), oob=rescale_none)


```


Ordered Root Length Bar Graph Relative to Control[trial]
```{r} 
library(scales)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtlngthtrial=="NA"))
yield<-( ddply(Y3, c("Title", "Var"), summarise,
               N    = length(rtlngthtrial),
               mean = mean(rtlngthtrial, na.rm=TRUE),
               sd   = sd(rtlngthtrial),
               se   = sd / sqrt(N),
               n=1))
limits<-aes(ymax=mean+se, ymin=mean-se)
y <- ggplot(yield, aes(y=mean,
                    x=factor(Title)))
y+geom_bar(position="dodge", aes(x=reorder(Title, -mean)),  stat="identity")+
facet_grid(Var~.)+
geom_errorbar(limits, position = "dodge")+
labs(title="Root Length Effect of Pythium spp. Inoculation", x="Pythium spp.", y="Root Length \n (% of the NTC, Rice)")+
  theme(plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=1), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 10),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(0,125), oob=rescale_none)


```


Ordered Shoot Wght Bar Graph Relative to Control[trial]
```{r} 
library(scales)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$shtwghttrial=="NA"))
yield<-( ddply(Y3, c("Title", "Var"), summarise,
               N    = length(shtwghttrial),
               mean = mean(shtwghttrial, na.rm=TRUE),
               sd   = sd(shtwghttrial),
               se   = sd / sqrt(N),
               n=1))
limits<-aes(ymax=mean+se, ymin=mean-se)
y <- ggplot(yield, aes(y=mean,
                    x=factor(Title)))
y+geom_bar(position="dodge", aes(x=reorder(Title, -mean)),  stat="identity")+
facet_grid(Var~.)+
geom_errorbar(limits, position = "dodge")+
labs(title="Shoot Weight Effect of Pythium spp. Inoculation", x="Pythium spp.", y="Shoot Weight \n (% of the NTC, Rice)")+
  theme(plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=1), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 10),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(0,125), oob=rescale_none)
```


Ordered Root Wght Bar Graph Relative to Control[trial]
```{r} 
library(scales)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA"))
yield<-( ddply(Y3, c("Title", "Var"), summarise,
               N    = length(rtwghttrial),
               mean = mean(rtwghttrial, na.rm=TRUE),
               sd   = sd(rtwghttrial),
               se   = sd / sqrt(N),
               n=1))
limits<-aes(ymax=mean+se, ymin=mean-se)
y <- ggplot(yield, aes(y=mean,
                    x=factor(Title)))
y+geom_bar(position="dodge", 
      aes(x=reorder(Title, -mean)), 
      stat="identity")+
facet_grid(Var~.)+
geom_errorbar(limits, position = "dodge")+
labs(title="Root Weight Effect of Pythium spp. Inoculation", x="Pythium spp.", y="Root Weight \n (% of the NTC, Rice)")+
  theme(plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=0.5), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 10),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(0,125), oob=rescale_none)
```


```{r}
library(scales)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
RHAllData<-subset(Y,!Y$Var=="Z")
boxplot(rtareatrial~Clade, data=RHAllData, las=2)
rb <- boxplot(rtareatrial~Clade, data=RHAllData, las=2)
mn.t <- tapply(RHAllData$rtareatrial, RHAllData$Clade, mean)
sd.t <- tapply(RHAllData$rtareatrial, RHAllData$Clade, sd)
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))
serror <- aggregate (RHAllData$rtareatrial, list(RHAllData$Clade), se)
serror <- tapply (RHAllData$rtareatrial, list(RHAllData$Clade), se)
xi <- seq(rb$n)
#optional
#optional
arrows(xi, mn.t - serror, xi, mn.t + serror,code = 3, col = "pink", angle = 75, length = .1)
points(xi, mn.t, col = "orange", pch = 18)
```

#setwd 
#This means that you can set a working directory and quite simply extract data from them!

You can have RMD with different formats - when making a new RMarkdown, you can select something "from template"
Use plain address as a link; if you want it to show up, you can use a ! before hand; [actual text] (call the file)

You can control output of Rmd files;{r chunk_name, ...}
cache, fig.width, fig.height

knit("markdownfilename.Rmd", "markdownfile.R")



par(mfrow=c(1,1),xpd=FALSE, mar=c(12,5,2.5,2+0.1))
RHrtm <- aggregate (RHAllData$emergetrial, list(RHAllData$Title), mean)
colnames(RHrtm) <-(c("Title", "emergetrial"))
plot(RHrtm$emergetrial, cex=0.6, las=2, xaxt="n", xlab="", ylab="", ylim=c(0,120))
axis(1, at=1:23, labels=RHrtm$Title, las=2, cex.axis=0.8)
mtext(Seedling Emergence", side=2, line=3.5)
mtext("As a Percentage of NTC, Rice", side=2, line=2.5, cex=0.63)
abline(h=100, lty=2, col="grey")
abline(h=0)
arrows (1:23, RHrtm$emergetrial - 1.96*serror$x, 1:23, RHrtm$emergetrial + 1.96*serror$x, code = 3, col = "red", angle = 90, length = .1)





```


