#High Resolution images, below:
#http://www.r-bloggers.com/high-resolution-figures-in-r/


#Agglomerative, Hierarchical Clustering
#Tests based on Lsmeans, Chord for RedHawk Cup Assay
```{r}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(cluster)
library(nlme)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA") & (Y$Var=="RH")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fitwght <- lmer(rtwght ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitarea <- lmer(rtarea ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitlngth <- lmer(rtlngth ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitemerge <- lmer(Emerge ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
ls1<-lsmeans(fitwght, contrast(trt.vs.ctrl, fitwght@frame$Title, ref=3)  ~ Title)
ls2<-lsmeans(fitarea, contrast(trt.vs.ctrl, fitarea@frame$Title, ref=3)  ~ Title)
ls3<-lsmeans(fitlngth, contrast(trt.vs.ctrl, fitlngth@frame$Title, ref=3)  ~ Title)
ls4<-lsmeans(fitemerge, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
wght1<-summary(ls1)
area1<-summary(ls2)
lngth1<-summary(ls3)
emerge1<-summary(ls4)
Dend<-cbind(wght1$lsmeans$lsmean, area1$lsmeans$lsmean,lngth1$lsmeans$lsmean,emerge1$lsmeans$lsmean)
Dend<-scale(as.matrix(Dend)) 
colnames(Dend)<-c("Root Dry Weight", "Root Area", "Root Length", "Emergence")
rownames(Dend)=wght1$lsmeans$Title
chord = function (s,diag) {
  # G.Nesslage, B.Maurer
  # computes chord distance
  # diag == TRUE returns full distance matrix
  # diag == FALSE returns distance matrix as dist object
  n = nrow(s)
  p <- ncol(s)
  D = matrix(0,n,n)
  m <- 1
  
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      nd <- matrix(0,p,1)
      dd1 <- matrix(0,p,1)
      dd2 <- matrix(0,p,1)
      for (k in 1:p){
        nd[k,]<- s[i,k]*s[j,k]
        dd1[k,] <- (s[i,k])^2
        dd2[k,] <- (s[j,k])^2
      }
      D[i,j] <- sqrt(2*(1-(sum(nd)/sqrt(sum(dd1)*
                                          sum(dd2)))))
      D[j,i] <- D[i,j]
    }
  }
  
  if(diag==FALSE) {D=as.dist(D)}
  return (D)
}

chord1 <- chord(Dend,diag=FALSE)
chord.v <- as.vector(chord1)
label <- rownames(Dend)
chord.dis <- agnes(chord1, method="average", diss = TRUE) 
chord.disv<-as.vector(chord.dis)
# very simple dendrogram
plot(chord.dis, ask = FALSE, which.plots = 2,labels = label)

#Cophenetic Distance and Correlation
chord.ave.co <- cophenetic(chord.dis)
chord.ave.cov <- as.vector(chord.ave.co)
cor(chord.ave.cov, chord.v)
plot(y = chord.ave.cov, x = chord.v)

#RMSE
rmse <- function(x, dis.matrix){
 r.m.s.e <- 1- (mean(as.vector(x)^2))/var(as.vector(dis.matrix))
 return(r.m.s.e)
}
chord.co.resid <- chord.ave.cov - chord.v
rmse(chord.co.resid, chord.v)

hc.rows<-hclust(chord.dis$diss)
jpeg("pheatmapRH.jpeg", width = 4, height = 4.75, units = 'in', res = 300)
pheatmap(Dend,  color = colorRampPalette(brewer.pal(n = 7, name =
"RdYlBu"))(100), show_rownames = TRUE, show_colnames = TRUE,
         cluster_rows=hc.rows,
         clustering_method = "average",
fontsize_row=8)
dev.off()

a<-c(expression(paste(italic("Y3$Title"), "Go")), expression(paste(italic("Y3$Title"), "Go")))


```

#Agglomerative, Hierarchical Clustering
#Tests based on Lsmeans, Chord for Zorro Cup Assay
```{r}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(cluster)
library(nlme)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA") & (Y$Var=="Z")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fitwght <- lmer(rtwght ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitarea <- lmer(rtarea ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitlngth <- lmer(rtlngth ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitemerge <- lmer(Emerge ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
ls1<-lsmeans(fitwght, contrast(trt.vs.ctrl, fitwght@frame$Title, ref=3)  ~ Title)
ls2<-lsmeans(fitarea, contrast(trt.vs.ctrl, fitarea@frame$Title, ref=3)  ~ Title)
ls3<-lsmeans(fitlngth, contrast(trt.vs.ctrl, fitlngth@frame$Title, ref=3)  ~ Title)
ls4<-lsmeans(fitemerge, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
wght1<-summary(ls1)
area1<-summary(ls2)
lngth1<-summary(ls3)
emerge1<-summary(ls4)
Dend<-cbind(wght1$lsmeans$lsmean, area1$lsmeans$lsmean,lngth1$lsmeans$lsmean,emerge1$lsmeans$lsmean)
Dend<-scale(as.matrix(Dend)) 
colnames(Dend)<-c("Root Dry Weight", "Root Area", "Root Length", "Emergence")
rownames(Dend)=wght1$lsmeans$Title
chord = function (s,diag) {
  # G.Nesslage, B.Maurer
  # computes chord distance
  # diag == TRUE returns full distance matrix
  # diag == FALSE returns distance matrix as dist object
  n = nrow(s)
  p <- ncol(s)
  D = matrix(0,n,n)
  m <- 1
  
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      nd <- matrix(0,p,1)
      dd1 <- matrix(0,p,1)
      dd2 <- matrix(0,p,1)
      for (k in 1:p){
        nd[k,]<- s[i,k]*s[j,k]
        dd1[k,] <- (s[i,k])^2
        dd2[k,] <- (s[j,k])^2
      }
      D[i,j] <- sqrt(2*(1-(sum(nd)/sqrt(sum(dd1)*
                                          sum(dd2)))))
      D[j,i] <- D[i,j]
    }
  }
  
  if(diag==FALSE) {D=as.dist(D)}
  return (D)
}

chord1 <- chord(Dend,diag=FALSE)
chord.v <- as.vector(chord1)
label <- rownames(Dend)
chord.dis <- agnes(chord1, method="average", diss = TRUE) 

hc.rows<-hclust(chord.dis$diss)
jpeg("pheatmapZ.jpeg", width = 4, height = 4.75, units = 'in', res = 300)
pheatmap(Dend,  color = colorRampPalette(brewer.pal(n = 7, name =
"RdYlBu"))(100), show_rownames = TRUE, show_colnames = TRUE,
         cluster_rows=hc.rows,
         clustering_method = "average",
fontsize_row=8)
dev.off()
```


Clade Disease
```{r, echo=FALSE}
library(scales)
library(plyr)
library(vegan)
library(cluster)
library(ade4)
library(stats)
library(tabplot)
library(ggplot2)
#update.packages("tabplot")
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y3<-subset(Y, !(Y$Sev=="NA") & !(Y$Clade=="Control")
)
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))
Y4<-aggregate(Y3$Sev, list(Y3$Clade, Y3$Temp), na.rm=TRUE, mean)
serror <- (aggregate(Y3$Sev, list(Y3$Clade, Y3$Temp), se))
Y4 <- cbind(Y4, serror$x)
colnames(Y4) <- c("Clade", "Temp", "sev", "SE")
limits<-aes(ymax=sev+(SE), ymin=sev-(SE))
Y4$Temp<-as.factor(Y4$Temp)

jpeg("IsoCladeSev.jpeg", width = 5, height = 8, units = 'in', res = 300)
y <- ggplot(Y4, aes(y=sev, 
                    x=as.factor(Clade), shape=Temp, color=Clade))
y+geom_point(aes(x=reorder(Clade, -sev)),  stat="identity", position=position_dodge(width=0.5), size=2)+
  theme_light()+
geom_errorbar(limits, stat="identity",position=position_dodge(width=0.5), width=0.4)+
labs(title="Seed Assay", x=expression(italic(Pythium)~Clade), y="Disease Severity Index")+
 theme(plot.title=element_text(size=18, colour="black"),
axis.title=element_text(size=11, colour="black"), axis.text.x=element_text(size=7, color="black", angle=90, hjust=1, vjust=0.5), axis.text.y=element_text(size=10, angle=90, hjust=0.5, vjust=0.5, color="black"))+
scale_y_continuous(limits=c(0,1), oob=rescale_none)+
  scale_x_discrete(breaks=Y4$Clade, labels=Y4$Clade)
  #geom_vline(xintercept=37.5)
dev.off()
```


#Emergence Across Varieties
```{r, echo=FALSE}
library(scales)
library(plyr)
library(lme4)
library(nlme)
library(lsmeans)
library(cluster)
library(ade4)
library(stats)
library(tabplot)
library(ggplot2)
#update.packages("tabplot")
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtareatrial=="NA") & !(Y$rtwghttrial=="NA"))
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))

Emerge <- lmer(emergetrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsEmerge<-lsmeans(Emerge, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
Emergelsmeans<-summary(lsEmerge)
v1<-c("Emergence %")
v2<-rep(v1, 31)
v3<-as.data.frame(Emergelsmeans$lsmeans$Title)
YEmerge <- as.data.frame(cbind(v3, Emergelsmeans$lsmeans$lsmean, Emergelsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$emergetrial, list(Y3$Title, Y3$Title), se)
YEmerge<-cbind(YEmerge, serror$x)
colnames(YEmerge) <- c("Title", "mean" , "SE", "Data", "se")

limits<-aes(ymax=mean+(se), ymin=mean-(se))
jpeg("Emerge.jpeg", width = 8.5, height = 4.5, units = 'in', res = 300)

y <- ggplot(YEmerge, aes(y=mean, 
                    x=Title))
y+geom_bar(aes(x=reorder(Title, -mean), fill=mean+(SE*1.96)>82), stat="identity", width=0.7)+
  geom_errorbar(limits, stat="identity", width=0.4)+
  theme_light()+
labs(title="Emergence", x="Pythium spp.", y="Emergence (%)")+
 theme(text=element_text(family="Arial Narrow"),
       plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=0.5), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 7, color="black"),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(-10,130), oob=rescale_none)
dev.off()

```



#Root Growth Parameters, Across Varieties
```{r, echo=FALSE}
library(scales)
library(plyr)
library(lme4)
library(nlme)
library(lsmeans)
library(cluster)
library(ade4)
library(stats)
library(tabplot)
library(ggplot2)
#update.packages("tabplot")
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtareatrial=="NA") & !(Y$rtwghttrial=="NA"))
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))

rtwght <- lmer(rtwghttrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtwght<-lsmeans(rtwght, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
rtwghtlsmeans<-summary(lsrtwght)
v1<-c("Root Weight %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtwghtlsmeans$lsmeans$Title)
Yrtwght <- as.data.frame(cbind(v3, rtwghtlsmeans$lsmeans$lsmean, rtwghtlsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$rtwghttrial, list(Y3$Title, Y3$Title), se)
Yrtwght<-cbind(Yrtwght, serror$x)
colnames(Yrtwght) <- c("Title", "mean" , "SE", "Data", "se")

rtarea <- lmer(rtareatrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtarea<-lsmeans(rtarea, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
rtarealsmeans<-summary(lsrtarea)
v1<-c("Root Area %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtarealsmeans$lsmeans$Title)
Yrtarea <- as.data.frame(cbind(v3, rtarealsmeans$lsmeans$lsmean, rtarealsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$rtareatrial, list(Y3$Title, Y3$Title), se)
Yrtarea<-cbind(Yrtarea, serror$x)
colnames(Yrtarea) <- c("Title", "mean" , "SE", "Data", "se")

rtlngth <- lmer(rtlngthtrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtlngth<-lsmeans(rtlngth, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
rtlngthlsmeans<-summary(lsrtlngth)
v1<-c("Root Length %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtarealsmeans$lsmeans$Title)
Yrtlngth <- as.data.frame(cbind(v3, rtlngthlsmeans$lsmeans$lsmean, rtlngthlsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$rtlngthtrial, list(Y3$Title, Y3$Title), se)
Yrtlngth<-cbind(Yrtlngth, serror$x)
colnames(Yrtlngth) <- c("Title", "mean" , "SE", "Data", "se")

Y4<-as.data.frame(rbind(Yrtwght, Yrtarea, Yrtlngth))

limits<-aes(ymax=mean+(se), ymin=mean-(se))
jpeg("Rootgrowth.jpeg", width = 8.5, height = 4.5, units = 'in', res = 300)

y <- ggplot(Y4, aes(y=mean, 
                    x=reorder(Title, -mean)))
y+geom_bar(data=Yrtwght, aes(x=reorder(Title, -mean), fill=mean>75.5), stat="identity", width=0.7)+
  geom_bar(data=Yrtlngth, aes(x=reorder(Title, -mean), fill=mean+(SE*1.96)>82), stat="identity", width=0.7)+
    geom_bar(data=Yrtarea, aes( x=reorder(Title, -mean), fill=mean+(SE*1.96)>82), stat="identity", width=0.7)+
  geom_errorbar(limits, stat="identity", width=0.4)+
  facet_grid(Data~.)+
  theme_light()+
labs(title="Seedling Assay", x="Pythium spp.", y="Percent Relative to NIC")+
 theme(text=element_text(family="Arial Narrow"),
       plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=0.5), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 7, color="black"),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(-10,130), oob=rescale_none)
dev.off()
```



#Graph of Cup Assay Parameters, Across Varieties #NO LONGER USED
```{r, echo=FALSE}
library(scales)
library(plyr)
library(lme4)
library(nlme)
library(lsmeans)
library(cluster)
library(ade4)
library(stats)
library(tabplot)
library(ggplot2)
#update.packages("tabplot")
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtareatrial=="NA") & !(Y$rtwghttrial=="NA"))
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))

rtwght <- lmer(rtwghttrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtwght<-lsmeans(rtwght, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
rtwghtlsmeans<-summary(lsrtwght)
v1<-c("Root Weight %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtwghtlsmeans$lsmeans$Title)
Yrtwght <- as.data.frame(cbind(v3, rtwghtlsmeans$lsmeans$lsmean, rtwghtlsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$rtwghttrial, list(Y3$Title, Y3$Title), se)
Yrtwght<-cbind(Yrtwght, serror$x)
colnames(Yrtwght) <- c("Title", "mean" , "SE", "Data", "se")

rtarea <- lmer(rtareatrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtarea<-lsmeans(rtarea, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
rtarealsmeans<-summary(lsrtarea)
v1<-c("Root Area %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtarealsmeans$lsmeans$Title)
Yrtarea <- as.data.frame(cbind(v3, rtarealsmeans$lsmeans$lsmean, rtarealsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$rtareatrial, list(Y3$Title, Y3$Title), se)
Yrtarea<-cbind(Yrtarea, serror$x)
colnames(Yrtarea) <- c("Title", "mean" , "SE", "Data", "se")

rtlngth <- lmer(rtlngthtrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtlngth<-lsmeans(rtlngth, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
rtlngthlsmeans<-summary(lsrtlngth)
v1<-c("Root Length %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtarealsmeans$lsmeans$Title)
Yrtlngth <- as.data.frame(cbind(v3, rtlngthlsmeans$lsmeans$lsmean, rtlngthlsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$rtlngthtrial, list(Y3$Title, Y3$Title), se)
Yrtlngth<-cbind(Yrtlngth, serror$x)
colnames(Yrtlngth) <- c("Title", "mean" , "SE", "Data", "se")

Emerge <- lmer(emergetrial ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsEmerge<-lsmeans(Emerge, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title)
Emergelsmeans<-summary(lsEmerge)
v1<-c("Emergence %")
v2<-rep(v1, 31)
v3<-as.data.frame(rtarealsmeans$lsmeans$Title)
YEmerge <- as.data.frame(cbind(v3, Emergelsmeans$lsmeans$lsmean, Emergelsmeans$lsmeans$SE, v2))
serror<-aggregate(Y3$Emergetrial, list(Y3$Title, Y3$Title), se)
YEmerge<-cbind(YEmerge, serror$x)
colnames(YEmerge) <- c("Title", "mean" , "SE", "Data", "se")

Y4<-as.data.frame(rbind(YEmerge, Yrtwght, Yrtarea, Yrtlngth))

limits<-aes(ymax=mean+(se), ymin=mean-(se))
jpeg("CupAssay.jpeg", width = 8.5, height = 4.5, units = 'in', res = 300)

y <- ggplot(Y4, aes(y=mean, 
                    x=Title))
y+geom_point(aes(x=reorder(Title, -mean), color=mean+(SE*1.96)<82), stat="identity", size=1.5)+
  geom_errorbar(limits, stat="identity", width=0.4)+
  facet_grid(Data~.)+
  theme_light()+
labs(title="", x="", y="Percent Relative to NIC")+
 theme(text=element_text(family="Arial Narrow"),
       plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=11, colour="black"), axis.text.x=element_text(size=9.5, color="black", face="italic", angle=90, hjust=1, vjust=0.5), axis.text.y=element_text(size=9.5, color="black"), strip.text.y = element_text(size = 7, color="black"),
legend.text=element_text(size=1),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(-10,130), oob=rescale_none)
dev.off()

```



Root Dry Weight
```{r, echo=FALSE}
library(scales)
library(plyr)
library(lme4)
library(nlme)
library(lsmeans)
library(cluster)
library(ade4)
library(stats)
library(tabplot)
library(ggplot2)
library(devtools)
#install_github("MarcoDVisser/choosecolor")
#library(choosecolor)
#update.packages("tabplot")
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtareatrial=="NA") & !(Y$rtwghttrial=="NA"))
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))
serror<-aggregate(Y3$rtwght, list(Y3$Title, Y3$Var), se)

rtwght <- lmer(rtwght ~ Title*Var + (1|Iso/Title:Trial:Set), data=Y3)
lsrtwght<-lsmeans(rtwght, pairwise~ Title | Var)
rtwghtlsmeans<-summary(lsrtwght)
v3<-as.data.frame(rtwghtlsmeans$lsmeans)
Yrtwght <- as.data.frame(cbind(v3))
RHlsmeans <- Yrtwght[-c(32:62),]
Zlsmeans <- Yrtwght[-c(1:31),]
limits<-aes(ymax=lsmean+(SE*1.96), ymin=lsmean-(SE*1.96))
limits2<-aes(ymax=lsmean+serror$x, ymin=lsmean-serror$x)
#R Colors
#http://research.stowers-institute.org/efg/R/Color/Chart/ColorChart.pdf
col1<-c("#0000FF", "#EE3B3B")

jpeg("RtWght.jpeg", width = 8.5, height = 4.5, units = 'in', res = 300)
y <- ggplot(Yrtwght, aes(y=lsmean, 
                    x=reorder(Title, -lsmean)))
y+geom_bar(data=RHlsmeans, aes(fill=(lsmean>.55)), stat="identity", width=0.7)+
  geom_bar(data=Zlsmeans, aes(fill=(lsmean>.30)), stat="identity", width=0.7)+
  geom_errorbar(limits2, stat="identity", width=0.4)+
  facet_grid(Var~.)+
  theme_light()+
labs(title="Root Dry Weight", x="Pythium spp.", y="Root Dry Weight (g)")+
 theme(text=element_text(family="Arial Narrow"),
       plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=9, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=0.5, face="italic"), axis.text.y=element_text(size=8, color="black"), strip.text.y = element_text(size = 7, color="black"),
strip.text.x = element_text(size = 9, color="black"))+ scale_y_continuous(limits=c(-0.05,0.8), oob=rescale_none)
#scale_color_manual(values = c("#0000FF", "#EE3B3B"))
dev.off()

```


#Sev Graph, Ordered (By Temp)
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y3<-subset(Y, !(Y$Sev=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
Y3$Temp<-as.character(Y3$Temp)

Y4<-( ddply(Y3, c("Title", "Temp"), summarise,
               N    = length(Sev),
            Sevsd = sd(Sev, na.rm=TRUE),
            Sevse = Sevsd / sqrt(N)))
fit3 <- lmer(Sev ~ Title*Var*Temp 
             + (1|Iso/(Title:Trial:Set)), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Temp)
lsmean1<-summary(ls1)
lsmean1<-as.data.frame(lsmean1$lsmeans)
#print(lsmean1)

limits<-aes(ymin=lsmean1$lsmean-(Y4$Sevse), ymax=lsmean1$lsmean+(Y4$Sevse))

limits2<-aes(ymin=lsmean1$lsmean-(1.96*lsmean1$SE), ymax=lsmean1$lsmean+(1.96*lsmean1$SE))
jpeg("SevbyTemp.jpeg", width = 5, height = 4, units = 'in', res = 300)
g<-ggplot(lsmean1, aes(x=Title, y=lsmean, shape=Temp, color=lsmean-(SE*1.96)>0.08))
g+ geom_point(aes(x=reorder(Title, -lsmean), shape=Temp), position=position_dodge(width=0.5), size=1)+
  theme_light()+
  #facet_grid(Var~.)+
geom_errorbar(limits, width=0.5, position=position_dodge(width=0.5))+
  labs(y="Disease Severity Index", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5, color="black", size=6, face="italic"),
        axis.title.y=element_text(size=7), legend.text=element_text(angle=0, color="black", size=1), legend.text.align=0.5, legend.title=element_text(angle=0, size=1), legend.title.align=0.5, legend.position="right", axis.text.y=element_text(angle=90, hjust=0.5, vjust=1, color="black", size=6), legend.text.align=0.5) + scale_colour_manual(name = 'p<0.05', values = setNames(c('red','dark green'), c(T, F))) +
  scale_y_continuous("Disease Severity Index", breaks=c(0, 0.5, 1), limits=c(-0.04, 1.03))
dev.off()
```


#Assay Comparison
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
library(graphics)
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/Assay.csv"))

jpeg("AssayCompare.jpeg", width = 8, height = 6, units = 'in', res = 300)
par(mfrow=c(2,4))
plot(Y$Emerge, (Y$Sev20), ylab="Disease Severity Index (20 C)", xlab="Emergence %")
plot(Y$RDW, (Y$Sev20), ylab="Disease Severity Index (20 C)", xlab="Root Dry Weight (g)")
plot(Y$RA, (Y$Sev20), ylab="Disease Severity Index (20 C)", xlab=expression(paste("Root Area (", cm^2, ")")))
plot(Y$RL, (Y$Sev20), ylab="Disease Severity Index (20 C)", xlab="Root Length (cm)")
plot(Y$Emerge, (Y$Sev26), ylab="Disease Severity Index (26 C)", xlab="Emergence %")
plot(Y$RDW, (Y$Sev26), ylab="Disease Severity Index (26 C)", xlab="Root Dry Weight (g)")
plot(Y$RA, (Y$Sev26), ylab="Disease Severity Index (26 C)", xlab=expression(paste("Root Area (", cm^2, ")")))
plot(Y$RL, (Y$Sev26), ylab="Disease Severity Index (26 C)", xlab="Root Length (cm)")
cor(Y$RDW, (Y$Sev20), method="pearson")
cor(Y$RA, (Y$Sev20), method="pearson")
cor(Y$RL, (Y$Sev20), method="pearson")
cor(Y$Emerge, (Y$Sev20), method="pearson")
cor(Y$RDW, (Y$Sev26), method="pearson")
cor(Y$RA, (Y$Sev26), method="pearson")
cor(Y$RL, (Y$Sev26), method="pearson")
cor(Y$Emerge, (Y$Sev26), method="pearson")
dev.off()
```




#Sev Graph, Ordered (By Temp*Var)
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y3<-subset(Y, !(Y$Sev=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
Y3$Temp<-as.character(Y3$Temp)

Y4<-( ddply(Y3, c("Title", "Var", "Temp"), summarise,
               N    = length(Sev),
            lsmean = mean(Sev, na.rm=TRUE),
            Sevsd = sd(Sev, na.rm=TRUE),
            Sevse = Sevsd / sqrt(N)))
fit3 <- lmer(Sev ~ Title*Var*Temp 
             + (1|Iso/(Title:Trial:Set)), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Temp:Var)
lsmean1<-summary(ls1)
lsmean1<-as.data.frame(lsmean1$lsmeans)
#print(lsmean1)
limits<-aes(ymin=lsmean1$lsmean-(Y4$Sevse), ymax=lsmean1$lsmean+(Y4$Sevse))
limits2<-aes(ymin=lsmean1$lsmean-(1.96*lsmean1$SE), ymax=lsmean1$lsmean+(1.96*lsmean1$SE))
lsmeanRH<-lsmean1[-c(61:120),] 
lsmeanZ<-lsmean1[-c(1:60),] 
Y4RH<-Y4[-c(61:120),] 
Y4Z<-Y4[-c(1:60),] 
g<-ggplot(lsmean1, aes(x=Title, y=lsmean, shape=Temp))
g+ geom_point(data=lsmeanRH, aes(x=reorder(Title, -lsmean), shape=Temp, color=lsmean-(Y4RH$Sevse)>0.1), position=position_dodge(width=0.5), size=4)+
geom_point(data=lsmeanZ, aes(x=reorder(Title, -lsmean), shape=Temp, color=lsmean-(1.96*Y4Z$Sevse)>0.25), position=position_dodge(width=0.5), size=4)+
  theme_light()+
  facet_grid(Var~.)+
geom_errorbar(limits, width=0.3, position=position_dodge(width=0.5))+
  labs(y="Disease Severity Index", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5, color="black", size=18),
        axis.title.y=element_text(size=22), legend.text=element_text(angle=0, color="black", size=22), legend.text.align=0.5, legend.title=element_text(angle=0, size=24), legend.title.align=0.5, legend.position="right", axis.text.y=element_text(angle=90, hjust=0.5, vjust=1, color="black", size=18), legend.text.align=0.5) + scale_colour_manual(name = 'p<0.05', values = setNames(c('red','dark green'), c(T, F))) +
  scale_y_continuous("Mean DSI Scores", breaks=c(0, 0.5, 1), limits=c(-0.2, 1.22))

```


#Note: The Following is Merger of Two Assays; Results are interesting, but not used:
#Agglomerative, Hierarchical Clustering
#Tests based on Lsmeans, Chord for RedHawk Cup Assay
```{r}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(cluster)
library(nlme)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y1 <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y2 <- subset(Y1, !(SevPercent=="NA") & (Var=="RH")
)
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$Title== "NIC, No Rice") & !(Y$rtareatrial=="NA") & (Y$Var=="RH")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trt<-as.character(Y3$Trt)
Y2$Iso<-as.character(Y2$Iso)
Y2$Set<-as.character(Y2$Set)
Y2$Trt<-as.character(Y2$Trt)
Y2$Trial<-as.character(Y2$Trial)
fitwght <- lmer(rtwght ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitarea <- lmer(rtarea ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitlngth <- lmer(rtlngth ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitemerge <- lmer(Emerge ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitsev<- lmer(Sev ~ Title + (1|Iso/Title:Trial:Set), data=Y2)
ls1<-lsmeans(fitwght, contrast(trt.vs.ctrl, fitwght@frame$Title, ref=3)  ~ Title)
ls2<-lsmeans(fitarea, contrast(trt.vs.ctrl, fitarea@frame$Title, ref=3)  ~ Title)
ls3<-lsmeans(fitlngth, contrast(trt.vs.ctrl, fitlngth@frame$Title, ref=3)  ~ Title)
ls4<-lsmeans(fitemerge, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
ls5<-lsmeans(fitsev, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
wght1<-summary(ls1)
area1<-summary(ls2)
lngth1<-summary(ls3)
emerge1<-summary(ls4)
sev1<-summary(ls5)
Dend<-cbind(wght1$lsmeans$lsmean, area1$lsmeans$lsmean,lngth1$lsmeans$lsmean,emerge1$lsmeans$lsmean, sev1$lsmeans$lsmean)
Dend<-scale(as.matrix(Dend)) 
colnames(Dend)<-c("Root Dry Weight", "Root Area", "Root Length", "Emergence", "Sev")
rownames(Dend)=wght1$lsmeans$Title
chord = function (s,diag) {
  # G.Nesslage, B.Maurer
  # computes chord distance
  # diag == TRUE returns full distance matrix
  # diag == FALSE returns distance matrix as dist object
  n = nrow(s)
  p <- ncol(s)
  D = matrix(0,n,n)
  m <- 1
  
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      nd <- matrix(0,p,1)
      dd1 <- matrix(0,p,1)
      dd2 <- matrix(0,p,1)
      for (k in 1:p){
        nd[k,]<- s[i,k]*s[j,k]
        dd1[k,] <- (s[i,k])^2
        dd2[k,] <- (s[j,k])^2
      }
      D[i,j] <- sqrt(2*(1-(sum(nd)/sqrt(sum(dd1)*
                                          sum(dd2)))))
      D[j,i] <- D[i,j]
    }
  }
  
  if(diag==FALSE) {D=as.dist(D)}
  return (D)
}

chord1 <- chord(Dend,diag=FALSE)
chord.v <- as.vector(chord1)
label <- rownames(Dend)
chord.dis <- agnes(chord1, method="average", diss = TRUE) 
chord.disv<-as.vector(chord.dis)
# very simple dendrogram
plot(chord.dis, ask = FALSE, which.plots = 2,labels = label)

#Cophenetic Distance and Correlation
chord.ave.co <- cophenetic(chord.dis)
chord.ave.cov <- as.vector(chord.ave.co)
cor(chord.ave.cov, chord.v)
plot(y = chord.ave.cov, x = chord.v)

#RMSE
rmse <- function(x, dis.matrix){
 r.m.s.e <- 1- (mean(as.vector(x)^2))/var(as.vector(dis.matrix))
 return(r.m.s.e)
}
chord.co.resid <- chord.ave.cov - chord.v
rmse(chord.co.resid, chord.v)

hc.rows<-hclust(chord.dis$diss)
jpeg("pheatmapRH.jpeg", width = 4, height = 4.75, units = 'in', res = 300)
pheatmap(Dend,  color = colorRampPalette(brewer.pal(n = 7, name =
"RdYlBu"))(100), show_rownames = TRUE, show_colnames = TRUE,
         cluster_rows=hc.rows,
         clustering_method = "average",
fontsize_row=8)
dev.off()

a<-c(expression(paste(italic("Y3$Title"), "Go")), expression(paste(italic("Y3$Title"), "Go")))


```




#Agglomerative, Hierarchical Clustering
#Tests based on Lsmeans, Chord for Zorro Cup Assay
```{r}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(cluster)
library(nlme)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y1 <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y2 <- subset(Y1, !(SevPercent=="NA") & (Var=="Z")
)
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$Title== "NIC, No Rice") & !(Y$rtareatrial=="NA") & (Y$Var=="Z")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trt<-as.character(Y3$Trt)
Y2$Iso<-as.character(Y2$Iso)
Y2$Set<-as.character(Y2$Set)
Y2$Trt<-as.character(Y2$Trt)
Y2$Trial<-as.character(Y2$Trial)
fitwght <- lmer(rtwght ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitarea <- lmer(rtarea ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitlngth <- lmer(rtlngth ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitemerge <- lmer(Emerge ~ Title + (1|Iso/Title:Trial:Set), data=Y3)
fitsev<- lmer(Sev ~ Title + (1|Iso/Title:Trial:Set), data=Y2)
ls1<-lsmeans(fitwght, contrast(trt.vs.ctrl, fitwght@frame$Title, ref=3)  ~ Title)
ls2<-lsmeans(fitarea, contrast(trt.vs.ctrl, fitarea@frame$Title, ref=3)  ~ Title)
ls3<-lsmeans(fitlngth, contrast(trt.vs.ctrl, fitlngth@frame$Title, ref=3)  ~ Title)
ls4<-lsmeans(fitemerge, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
ls5<-lsmeans(fitsev, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
wght1<-summary(ls1)
area1<-summary(ls2)
lngth1<-summary(ls3)
emerge1<-summary(ls4)
sev1<-summary(ls5)
Dend<-cbind(wght1$lsmeans$lsmean, area1$lsmeans$lsmean,lngth1$lsmeans$lsmean,emerge1$lsmeans$lsmean, sev1$lsmeans$lsmean)
Dend<-scale(as.matrix(Dend)) 
colnames(Dend)<-c("Root Dry Weight", "Root Area", "Root Length", "Emergence", "Sev")
rownames(Dend)=wght1$lsmeans$Title
chord = function (s,diag) {
  # G.Nesslage, B.Maurer
  # computes chord distance
  # diag == TRUE returns full distance matrix
  # diag == FALSE returns distance matrix as dist object
  n = nrow(s)
  p <- ncol(s)
  D = matrix(0,n,n)
  m <- 1
  
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      nd <- matrix(0,p,1)
      dd1 <- matrix(0,p,1)
      dd2 <- matrix(0,p,1)
      for (k in 1:p){
        nd[k,]<- s[i,k]*s[j,k]
        dd1[k,] <- (s[i,k])^2
        dd2[k,] <- (s[j,k])^2
      }
      D[i,j] <- sqrt(2*(1-(sum(nd)/sqrt(sum(dd1)*
                                          sum(dd2)))))
      D[j,i] <- D[i,j]
    }
  }
  
  if(diag==FALSE) {D=as.dist(D)}
  return (D)
}

chord1 <- chord(Dend,diag=FALSE)
chord.v <- as.vector(chord1)
label <- rownames(Dend)
chord.dis <- agnes(chord1, method="average", diss = TRUE) 
chord.disv<-as.vector(chord.dis)
# very simple dendrogram
plot(chord.dis, ask = FALSE, which.plots = 2,labels = label)

#Cophenetic Distance and Correlation
chord.ave.co <- cophenetic(chord.dis)
chord.ave.cov <- as.vector(chord.ave.co)
cor(chord.ave.cov, chord.v)
plot(y = chord.ave.cov, x = chord.v)

#RMSE
rmse <- function(x, dis.matrix){
 r.m.s.e <- 1- (mean(as.vector(x)^2))/var(as.vector(dis.matrix))
 return(r.m.s.e)
}
chord.co.resid <- chord.ave.cov - chord.v
rmse(chord.co.resid, chord.v)

hc.rows<-hclust(chord.dis$diss)
jpeg("pheatmapZ.jpeg", width = 4, height = 4.75, units = 'in', res = 300)
pheatmap(Dend,  color = colorRampPalette(brewer.pal(n = 7, name =
"RdYlBu"))(100), show_rownames = TRUE, show_colnames = TRUE,
         cluster_rows=hc.rows,
         clustering_method = "average",
fontsize_row=8)
dev.off()

a<-c(expression(paste(italic("Y3$Title"), "Go")), expression(paste(italic("Y3$Title"), "Go")))


```


