---
title: "DryBeanJune10.2016"
author: "Devon"
date: "June 10, 2016"
output: html_document
---

#Agglomerative, Hierarchical Clustering
#Tests based on Lsmeans, Chord for RedHawk Cup Assay
```{r}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(cluster)
library(nlme)
library(ggplot2)
library(pheatmap)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA") & (Y$Var=="RH")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fitwght <- lmer(rtwght ~ Title + (1|Iso/Title/Set), data=Y3)
fitarea <- lmer(rtarea ~ Title + (1|Iso/Title/Set), data=Y3)
fitlngth <- lmer(rtlngth ~ Title + (1|Iso/Title/Set), data=Y3)
fitemerge <- lmer(Emerge ~ Title + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fitwght, contrast(trt.vs.ctrl, fitwght@frame$Title, ref=3)  ~ Title)
ls2<-lsmeans(fitarea, contrast(trt.vs.ctrl, fitarea@frame$Title, ref=3)  ~ Title)
ls3<-lsmeans(fitlngth, contrast(trt.vs.ctrl, fitlngth@frame$Title, ref=3)  ~ Title)
ls4<-lsmeans(fitemerge, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
wght1<-summary(ls1)
area1<-summary(ls2)
lngth1<-summary(ls3)
emerge1<-summary(ls4)
Dend<-cbind(wght1$lsmeans$lsmean, area1$lsmeans$lsmean,lngth1$lsmeans$lsmean,emerge1$lsmeans$lsmean)
Dend<-scale(as.matrix(Dend)) 
colnames(Dend)<-c("Root Dry Weight", "Root Area", "Root Lenght", "Emergence")
rownames(Dend)=wght1$lsmeans$Title
chord = function (s,diag) {
  # G.Nesslage, B.Maurer
  # computes chord distance
  # diag == TRUE returns full distance matrix
  # diag == FALSE returns distance matrix as dist object
  n = nrow(s)
  p <- ncol(s)
  D = matrix(0,n,n)
  m <- 1
  
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      nd <- matrix(0,p,1)
      dd1 <- matrix(0,p,1)
      dd2 <- matrix(0,p,1)
      for (k in 1:p){
        nd[k,]<- s[i,k]*s[j,k]
        dd1[k,] <- (s[i,k])^2
        dd2[k,] <- (s[j,k])^2
      }
      D[i,j] <- sqrt(2*(1-(sum(nd)/sqrt(sum(dd1)*
                                          sum(dd2)))))
      D[j,i] <- D[i,j]
    }
  }
  
  if(diag==FALSE) {D=as.dist(D)}
  return (D)
}

chord1 <- chord(Dend,diag=FALSE)
chord.v <- as.vector(chord1)
label <- rownames(Dend)
chord.dis <- agnes(chord1, method="average", diss = TRUE) 
chord.disv<-as.vector(chord.dis)
# very simple dendrogram
plot(chord.dis, ask = FALSE, which.plots = 2,labels = label)

#Cophenetic Distance and Correlation
chord.ave.co <- cophenetic(chord.dis)
chord.ave.cov <- as.vector(chord.ave.co)
cor(chord.ave.cov, chord.v)
plot(y = chord.ave.cov, x = chord.v)

#RMSE
rmse <- function(x, dis.matrix){
 r.m.s.e <- 1- (mean(as.vector(x)^2))/var(as.vector(dis.matrix))
 return(r.m.s.e)
}
chord.co.resid <- chord.ave.cov - chord.v
rmse(chord.co.resid, chord.v)

hc.rows<-hclust(chord.dis$diss)
pheatmap(Dend, show_rownames = TRUE, show_colnames = TRUE,
         cluster_rows=hc.rows,
         clustering_method = "average",
fontsize_row=8)
```

#Agglomerative, Hierarchical Clustering
#Tests based on Lsmeans, Chord for Zorro Cup Assay
```{r}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(cluster)
library(nlme)
library(ggplot2)
library(pheatmap)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA") & (Y$Var=="Z")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fitwght <- lmer(rtwght ~ Title + (1|Iso/Title/Set), data=Y3)
fitarea <- lmer(rtarea ~ Title + (1|Iso/Title/Set), data=Y3)
fitlngth <- lmer(rtlngth ~ Title + (1|Iso/Title/Set), data=Y3)
fitemerge <- lmer(Emerge ~ Title + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fitwght, contrast(trt.vs.ctrl, fitwght@frame$Title, ref=3)  ~ Title)
ls2<-lsmeans(fitarea, contrast(trt.vs.ctrl, fitarea@frame$Title, ref=3)  ~ Title)
ls3<-lsmeans(fitlngth, contrast(trt.vs.ctrl, fitlngth@frame$Title, ref=3)  ~ Title)
ls4<-lsmeans(fitemerge, contrast(trt.vs.ctrl, fitemerge@frame$Title, ref=3)  ~ Title)
wght1<-summary(ls1)
area1<-summary(ls2)
lngth1<-summary(ls3)
emerge1<-summary(ls4)
Dend<-cbind(wght1$lsmeans$lsmean, area1$lsmeans$lsmean,lngth1$lsmeans$lsmean,emerge1$lsmeans$lsmean)
Dend<-scale(as.matrix(Dend)) 
colnames(Dend)<-c("Root Dry Weight", "Root Area", "Root Length", "Emergence")
rownames(Dend)=wght1$lsmeans$Title
chord = function (s,diag) {
  # G.Nesslage, B.Maurer
  # computes chord distance
  # diag == TRUE returns full distance matrix
  # diag == FALSE returns distance matrix as dist object
  n = nrow(s)
  p <- ncol(s)
  D = matrix(0,n,n)
  m <- 1
  
  for (i in 1:(n-1)){
    for (j in (i+1):n){
      nd <- matrix(0,p,1)
      dd1 <- matrix(0,p,1)
      dd2 <- matrix(0,p,1)
      for (k in 1:p){
        nd[k,]<- s[i,k]*s[j,k]
        dd1[k,] <- (s[i,k])^2
        dd2[k,] <- (s[j,k])^2
      }
      D[i,j] <- sqrt(2*(1-(sum(nd)/sqrt(sum(dd1)*
                                          sum(dd2)))))
      D[j,i] <- D[i,j]
    }
  }
  
  if(diag==FALSE) {D=as.dist(D)}
  return (D)
}

chord1 <- chord(Dend,diag=FALSE)
chord.v <- as.vector(chord1)
label <- rownames(Dend)
chord.dis <- agnes(chord1, method="average", diss = TRUE) 
chord.disv<-as.vector(chord.dis)
# very simple dendrogram
plot(chord.dis, ask = FALSE, which.plots = 2,labels = label)

#RMSE
rmse <- function(x, dis.matrix){
 r.m.s.e <- 1- (mean(as.vector(x)^2))/var(as.vector(dis.matrix))
 return(r.m.s.e)
}
chord.co.resid <- chord.ave.cov - chord.v
rmse(chord.co.resid, chord.v)

hc.rows<-hclust(chord.dis$diss)
pheatmap(Dend, show_rownames = TRUE, show_colnames = TRUE,
         cluster_rows=hc.rows,
         clustering_method = "average",
fontsize_row=8)

```


Seed Rot Zorro, two temperatures
```{r}
library(pheatmap)
library(plyr)
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y20<-subset(Y, #!(Y$SevPercent=="NA") & 
              (Y$Var=="Z") & Y$Temp=="20")
Y26<-subset(Y, #!(Y$SevPercent=="NA") & 
              (Y$Var=="Z") & Y$Temp=="26")
Sev26<-Y26$SevPercent
Y3<-cbind(Y20, Sev26)
Y3$tisolate<-paste(Y3$Title,Y3$Trt)
Y4<-( ddply(Y3, c("Var", "Title"), summarise,
               N    = length(SevPercent),
            sev20 = mean(SevPercent, na.rm=TRUE),
            sev26 = mean(Sev26, na.rm=TRUE)))
stuff<-as.data.frame(cbind(Y4$sev20, Y4$sev26))
rownames(stuff)<-Y4$Title
colnames(stuff)<-c("Severity 20C", "Severity 26C")
pheatmap(stuff, show_rownames = TRUE, show_colnames = TRUE,
         #cluster_rows=hc.rows, 
         clustering_method="average")

```

Seed Rot Red Hawk, two temperatures
```{r}
library(pheatmap)
library(plyr)
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y20<-subset(Y, #!(Y$SevPercent=="NA") & 
              (Y$Var=="RH") & Y$Temp=="20")
Y26<-subset(Y, #!(Y$SevPercent=="NA") & 
              (Y$Var=="RH") & Y$Temp=="26")
Sev26<-Y26$SevPercent
Y3<-cbind(Y20, Sev26)
Y3$tisolate<-paste(Y3$Title,Y3$Trt)
Y4<-( ddply(Y3, c("Var", "Title"), summarise,
               N    = length(SevPercent),
            sev20 = mean(SevPercent, na.rm=TRUE),
            sev26 = mean(Sev26, na.rm=TRUE)))
stuff<-as.data.frame(cbind(Y4$sev20, Y4$sev26))
rownames(stuff)<-Y4$Title
colnames(stuff)<-c("Severity 20C", "Severity 26C")
pheatmap(stuff, show_rownames = TRUE, show_colnames = TRUE,
         #cluster_rows=hc.rows, 
         clustering_method="average")

```


#Virulence-sorted Severity by Temperature - Treatment
```{r, echo=FALSE}
library(scales)
library(plyr)
library(vegan)
library(cluster)
library(ade4)
library(stats)
library(tabplot)
library(ggplot2)
#update.packages("tabplot")
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y3<-subset(Y, !(Y$Sev=="NA") #& !(Y$Var=="Z"))
)
Y4<-aggregate(Y3$Sev, list(Y3$Trt, Y3$Clade, Y3$Temp), na.rm=TRUE, mean)
se <- function(x) sqrt(var(x, na.rm=TRUE)/(length(na.omit(x))))
serror <- (aggregate(Y3$Sev, list(Y3$Trt, Y3$Clade, Y3$Temp), se))
Y4 <- cbind(Y4, serror$x)
colnames(Y4) <- c("trt", "Clade", "Temp", "sev", "SE")
limits<-aes(ymax=sev+(1.96*SE), ymin=sev-(1.96*SE))
Y4$Temp<-as.factor(Y4$Temp)

y <- ggplot(Y4, aes(y=sev, 
                    #fill=sdtrt, 
                    x=factor(Y4$trt), shape=Y4$Temp, color=Y4$Clade))
y+geom_point(position="dodge", aes(x=reorder(trt, -sev)),  stat="identity", size=2.75)+
  theme_light()+
geom_errorbar(limits, stat="identity", width=0.3)+
labs(title="Seed Assay", x="Pythium spp.", y="Disease Severity Score")+
 theme(plot.title=element_text(size=20, colour="black"),
axis.title=element_text(size=13, colour="black"), axis.text.x=element_text(size=8, color="black", angle=90, hjust=1, vjust=1), axis.text.y=element_text(size=10, color="black"), strip.text.y = element_text(size = 10),
strip.text.x = element_text(size = 10, color="black"))+
scale_y_continuous(limits=c(0,1), oob=rescale_none)+
  geom_vline(xintercept=37.5)

```

#RtArea Graph, Ordered
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fit3 <- lmer(rtarea ~ Title*Var + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Var)
lsmean1<-summary(ls1)
print(lsmean1)
limits<-aes(ymin=lsmean-(1.96*SE), ymax=lsmean+(1.96*SE))
g<-ggplot(lsmean1$lsmeans, aes(x=Title, y=lsmean))
g+ geom_point(aes(x=reorder(Title, -lsmean)))+
  theme_light()+
  facet_grid(Var~.)+
geom_errorbar(limits, width=0.2)+labs(y="Mean Root Area\n(in square cm)", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5))
```

#RtLngth Graph, Ordered
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fit3 <- lmer(rtlngth ~ Title*Var + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Var)
lsmean1<-summary(ls1)
print(lsmean1)
limits<-aes(ymin=lsmean-(1.96*SE), ymax=lsmean+(1.96*SE))
g<-ggplot(lsmean1$lsmeans, aes(x=Title, y=lsmean))
g+ geom_point(aes(x=reorder(Title, -lsmean)))+
  theme_light()+
  facet_grid(Var~.)+
geom_errorbar(limits, width=0.2)+labs(y="Mean Root Length\n(in cm)", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5))
```


#RtWght Graph, Ordered
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fit3 <- lmer(rtwght ~ Title*Var + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Var)
lsmean1<-summary(ls1)
limits<-aes(ymin=lsmean-(1.96*SE), ymax=lsmean+(1.96*SE))
g<-ggplot(lsmean1$lsmeans, aes(x=Title, y=lsmean))
g+ geom_point(aes(x=reorder(Title, -lsmean), colour=(lsmean+1.96*SE)<0.4))+
  theme_light()+
  facet_grid(Var~.)+
geom_errorbar(limits, width=0.15, color="dark grey")+ylab(expression(paste('Mean Root Area ', (cm^{2})))) +
  #labs(y="Mean Root Area \n(in cm2)", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5))+
  scale_colour_manual(name = 'Significant', values = setNames(c('red','green'), c(T, F)))

```


#RtArea Graph, Ordered
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllCupApr19.csv"))
Y3<-subset(Y, !(Y$rtwghttrial=="NA") & !(Y$rtareatrial=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fit3 <- lmer(rtarea ~ Title*Var + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Var)
lsmean1<-summary(ls1)
limits<-aes(ymin=lsmean-(1.96*SE), ymax=lsmean+(1.96*SE))
g<-ggplot(lsmean1$lsmeans, aes(x=Title, y=lsmean))
g+ geom_point(aes(x=reorder(Title, -lsmean), colour=(lsmean+1.96*SE)<10.5))+
  theme_light()+
  facet_grid(Var~.)+
geom_errorbar(limits, width=0.15, color="dark grey")+ylab(expression(paste('Mean Root Area ', (cm^{2})))) +
  #labs(y="Mean Root Area \n(in cm2)", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5))+
  scale_colour_manual(name = 'Significant', values = setNames(c('red','green'), c(T, F)))
```


#Sev Graph, Ordered (By Var, Temp)
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
library(plyr)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y3<-subset(Y, !(Y$Sev=="NA")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
Y3$Temp<-as.character(Y3$Temp)

Y4<-( ddply(Y3, c("Title", "Temp"), summarise,
               N    = length(Sev),
            Sevsd = sd(Sev, na.rm=TRUE),
            Sevse = Sevsd / sqrt(N)))
fit3 <- lmer(Sev ~ Title*Var*Temp 
             + (1|Iso/Title/Set), data=Y3)
ls1<-lsmeans(fit3, contrast(trt.vs.ctrl, fit3@frame$Title, ref=3)  ~ Title:Temp)
lsmean1<-summary(ls1)
#print(lsmean1)
limits<-aes(ymin=lsmean1$lsmeans$lsmean-(1.96*Y4$Sevse), ymax=lsmean1$lsmeans$lsmean+(1.96*Y4$Sevse))
g<-ggplot(lsmean1$lsmeans, aes(x=Title, y=lsmean, shape=Temp, color=lsmean-(SE*1.96)>0.16))
g+ geom_point(aes(x=reorder(Title, -lsmean), shape=Temp), position=position_dodge(width=0.5), size=4)+
  theme_light()+
  #facet_grid(Var~.)+
geom_errorbar(limits, width=0.3, position=position_dodge(width=0.5))+
  labs(y="Disease Severity Index", x="")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5, color="black", size=18),
        axis.title.y=element_text(size=22), legend.text=element_text(angle=0, color="black", size=22), legend.text.align=0.5, legend.title=element_text(angle=0, size=24), legend.title.align=0.5, legend.position="right", axis.text.y=element_text(angle=90, hjust=0.5, vjust=1, color="black", size=18), legend.text.align=0.5) + scale_colour_manual(name = 'p<0.05', values = setNames(c('red','dark green'), c(T, F)))
```


Assay Comparison
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
library(ggplot2)
library(graphics)
rm(list = ls())
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/Assay.csv"))
par(mfrow=c(2,4))
plot(Y$Emerge, (Y$Sev20), ylab="Disease Severity Index \n 20 C", xlab="Emergence %")
plot(Y$RDW, (Y$Sev20), ylab="Disease Severity Index \n 20 C", xlab="Root Dry Weight \n (g)")
plot(Y$RA, (Y$Sev20), ylab="Disease Severity Index \n 20 C", xlab="Root Area \n (sq. cm)")
plot(Y$RL, (Y$Sev20), ylab="Disease Severity Index \n 20 C", xlab="Root Length \n (cm)")
plot(Y$Emerge, (Y$Sev26), ylab="Disease Severity Index \n 26 C", xlab="Emergence %")
plot(Y$RDW, (Y$Sev26), ylab="Disease Severity Index \n 26 C", xlab="Root Dry Weight \n (g)")
plot(Y$RA, (Y$Sev26), ylab="Disease Severity Index \n 26 C", xlab="Root Area \n (sq. cm)")
plot(Y$RL, (Y$Sev26), ylab="Disease Severity Index \n 26 C", xlab="Root Length \n (cm)")
cor(Y$RL, (Y$Sev20))
cor(Y$RA, (Y$Sev20))
cor(Y$RL, (Y$Sev20))
cor(Y$Emerge, (Y$Sev20))
cor(Y$RL, (Y$Sev26))
cor(Y$RA, (Y$Sev26))
cor(Y$RL, (Y$Sev26))
cor(Y$Emerge, (Y$Sev26))

```

Model Selection Disease Severity
```{r, echo=FALSE}
library(lme4)
library(lsmeans)
library(pbkrtest)
library(nlme)
Y <- (read.csv("/Users/devonrossman/Desktop/CSVforR/DryBean/AllRotMay9.csv"))
Y3<-subset(Y, !(Y$Sev=="NA") & #(Y$Var=="Z") & 
             !(Y$Trt=="2.5")
)
Y3$Trt<-as.character(Y3$Trt)
Y3$Iso<-as.character(Y3$Iso)
Y3$Set<-as.character(Y3$Set)
Y3$Trial<-as.character(Y3$Trial)
fit1 <- lmer(Sev ~ Title*Temp*Var - Title:Temp:Var + (1| Iso/(Title:Trial:Set)), data=Y3)
fit2 <- lmer(Sev ~ Title*Temp*Var - Title:Temp:Var + (1|Iso/Title/Set), data=Y3)
fit3 <- lmer(Sev ~ Title*Temp + (1|Iso/Title/Trial), data=Y3)
fit4 <- lmer(Sev ~ Title*Temp + (1|Iso/Title), data=Y3)
a<-anova(fit1,fit2)
print(a)





```